### 2013-08-21 proposal/proposal/chapter3.tex
Changes

\setcounter{chapter}{3}
\chapter{Contribution and Research Activities}

%******************** START Contribution and Research Activities. ***************
In this chapter we will introduce the important notions in argumentation-based negotiation,  
\section{Argumentation System and Agent Theory}\label{sec:Argumentation}
Let us recall some basic definitions in the argumentation-based negotiation.

\begin{definition}{\emph{[Move].}} \label{Move}
Let $Ag = \{Ag_1,Ag_2, \ldots\}$ be a set of symbols representing agent names that may be involved in a negotiation dialogue $D$. A
move $M_i\in D$ consists of the agent that utters the move: $Speaker(M_i)\in Ag$; the set of agents to which the move is
addressed: $Hearer(M_i)\subseteq Ag$; and the content of the move: $Content(M_i)$.
\end{definition}

For simplification reasons, we will consider bilateral negotiation (i.e., only two agents involved in the dialogue $Ag_1$ and $Ag_2$).
Multi-party dialogues is an important issue that we plan to investigate in future work. During a dialogue, several moves may be uttered.
Those moves constitute a dialogue $D$, which is a sequence of moves denoted by $[M_0$,$M_1$,\ldots,$M_n]$, where $M_0$ is the initial move,
$M_n$ is the final one, and $|D|$ is the length of the dialogue (i.e., $|D| = |[M_0$,$M_1$,\ldots,$M_n]| = n+1$).

In what follows, we assume that the agents exchange the content of the moves during the dialogue as arguments because our approach is
argumentation-based. In argumentation systems the logical language $\pounds$ is essential, we will recall the definition of the
argument concept, a definition of the attack relation between arguments, and the definition of acceptability
from~\cite{Amgoud00}, then we will use these definitions in our approach. Here $\Gamma$ indicates a possibly inconsistent
knowledge base with no deductive closure, and $\vdash$ stands for classical inference.

\begin{definition}{\emph{[Argument].}} \label{Argument}
An argument $Arg$ is a pair $(H, h)$ where $h$ is a formula of $\pounds$ and $H$ a subset of  $\Gamma$ such that: i) $H$ is
consistent, ii) $H \vdash h$ and iii) $H$ is minimal, so that no subset of $H$ satisfying both i and ii exists. $H$ is called the
support of the argument and $h$ its conclusion.
\end{definition}

\begin{definition}{\emph{[Attack].}} \label{Attack}
Let $(H, h)$, $(H^{\prime}, h^{\prime})$ be two arguments. $(H^{\prime}, h^{\prime})$ attacks $(H, h)$ iff $H^{\prime} \vdash
\neg h$. In other words, an argument is attacked iff there is an argument for the negation of its conclusion.
\end{definition}

Our dialogue architecture also involves general knowledge, such as knowledge about the dialogue subject. Agents can also reason about
their preferences in relation to beliefs. The idea is to capture the fact that some facts are believed more than others. For this
reason, we assume, like in~\cite{Mbarki06}, that any set of facts has a preference order over it. We suppose that this ordering
derives from the fact that the agent's knowledge base ($\Gamma$) is stratified into non-overlapping sets $\Gamma_1,\cdots ,
\Gamma_n$ such that facts in $\Gamma_i$ are all equally preferred and are more preferred than those in $\Gamma_j$ where $i < j$. The
preference level of a subset of $\Gamma$ whose elements belong to different non-overlapping sets is defined as follows.

\begin{definition}{\emph{[Preference Level].}} \label{preferenceLevel}
The preference level of a nonempty subset $\gamma$ of $\Gamma$ denoted by level $(\gamma)$ is the number of the highest numbered
layer which has a member in $\gamma$.
\end{definition}
\begin{example}
Let $\Gamma = \Gamma_1 \bigcup \Gamma_2 $ with $\Gamma_1$ = $\{ a, b\}$ and $\Gamma_2$ = $\{ c,d \}$ and $\gamma = \{a\}$ and
$\gamma\prime = \{a,d\}$. We have: level$(\gamma) = 1$ and level$(\gamma\prime) = 2$.
\end{example}

\subsection{Strategic and Tactic Reasoning}\label{sec:reasoning}
A preliminary framework for strategic and tactic reasoning for agent communication was proposed in~\cite{Mbarki06}. This
reasoning framework is specified using argumentation theory combined to a relevance theory. Strategic reasoning enables agents
to decide about the global communication plan in terms of the macro-actions to perform in order to achieve the main dialogue
goal. Tactic reasoning, on the other hand, allows agents to locally select, at each dialogue step the most appropriate
argument according to the adopted strategy.

In this paper, we refine, and use the notion of strategic and tactic reasoning. The agent uses his tactic reasoning at each dialogue
step to measure his uncertainty about selecting the right move to achieve some sub goals of the global goal (i.e., the agreement in our case).
In other words, how agent can select the right argument from a set of possible choices at certain dialogue step depends on his tactic by assigning
different probability values to the different arguments and select the one with higher probability to be accepted by his opponent. On the other hand,
this tactic is based on the adapted strategy that the agent built before starting the negotiation to achieve the agreement. Moreover, we study the
relationship between these two types of uncertainty not only when a unique move is considered, but also when the whole dialogue is being analyzed.

%\subsubsection{Strategic Reasoning}\label{sec:Strategic}
\textbf{Strategic Reasoning}\\
Before engaging in a negotiation, agents must build a global strategy on the sub-goals to achieve. Sub-goals determine the general steps to follow so
that the global goal can be realized. Strategy is subject to the agent's current beliefs and constraints, such as the agent's budget and negotiation
time limit. To achieve the same negotiation goal, an agent can have several alternative strategies reflected by different sets of subgoals. The dialogue
goal, sub-goals, and constraints can be expressed using propositional logic. The set of constraints can be inconsistent, but the sub-set of those constraints
and the sub-set of beliefs the agent decides to consider should be consistent. We define the strategy as a function that associates to a goal and a sub-set
of consistent beliefs and constraints a sub-set of alternatives, each of which is a set of sub goals, which means an element of the set $2^{2^{\mathbb{B}}}$,
where $\mathbb{B}$ is the set of goals.

\begin{definition}{\emph{[Strategy]\footnote{This definition is different than the one proposed in~\cite{Mbarki06}, in the sense that this function
associates a set of sub-goals to a sub-set of agent's knowledge base and constraints instead of associating a set of goals to a set of operational constraint
and a set of conversational criterions.}. }} \label{strategy}
Let $\mathbb{B}$ be a set of goals, $\mathbb{C}$ a set of constraints and $\Gamma$ the agent's knowledge base. A strategy is a function:
\begin{equation}\label{equation2-}
\mathcal{S}: \mathbb{B} \times 2^{\mathbb{C}} \times 2^{\Gamma}
\rightarrow 2^{2^{\mathbb{B}}}
\end{equation}
\end{definition}

%\subsubsection{Tactic Reasoning}\label{sec:tactic}
\textbf{Tactic Reasoning} \\
Tactics allow agents to select one action (i.e., the content of the move or the argument) from a set of possible actions (i.e., set of
arguments) in order to achieve a sub-goal as computed by the adapted strategy. The purpose of this theory is to ensure that the selected argument
is the most appropriate one according to the current context i.e., the less risk of failure, the more favorable and the more preferable one,
which means the move with the higher probability. Our aim is to investigate the uncertainty issues associated with the agent while taking decision
and selecting the arguments. By refining some definition in the work presented in~\cite{Mbarki06} such as the definition of strategy, context,
tactic, and risk of failure. We will advocate some criteria that help to compute and analyze the agents' uncertainty at each negotiation step i.e., at
the tactic level. and based on these criteria we classify the set of possible arguments into three different classes ``\textit{Arguments Classification}",
which will be introduced later on in section~\ref{sec:classification}. This also will help the agents to assign the probability to each possible argument
and use it in measuring the uncertainty using Shannon Entropy as it will be explained in section~\ref{moveuncertainty}. Arguments classification is based
on the relevance of arguments, risk of failure, favorite and preference relations. In the following we will discuss the definitions of these notions.

\textbf{Relevance of Argument.} \\
We define the relevance of argument as in~\cite{Mbarki06} according to the conversation context. This will allow the participants in the dialogue to
select the most relevant argument from the set of possible arguments at certain dialogue step by taking into account the last communicative act as well as the
previous ones. This will give the agents the ability of backtracking in case the choice is shown to be incorrect because the selected argument is not accepted
by the addressee. The arguments are ordered based on their relevance, and the process of selecting arguments is called \emph{Arguments Selection Mechanism}.

%The authors in~\cite{Mbarki06}, have defined the relevance of an argument according to the conversation context. The objective is
%to allow agents to select the most relevant argument from the set of possible arguments at a given dialogue step by taking into
%account the last communicative act as well as the previous ones. This will give the participants the ability of backtracking in
%case the choice is shown to be incorrect because the selected argument is not accepted by the addressee. The arguments are
%ordered based on their relevance, and the process of selecting arguments is called \emph{Arguments Selection Mechanism}.

\textbf{Arguments Selection Mechanism.} \\
Let $\pounds$ be a logical language. We define the conversation context for an agent $Ag_1$ committed in a conversation with
another agent $Ag_2$ as follows:

\begin{definition}{\emph{[Context].}} \label{context}
The conversation context for an agent $Ag_1$ (the speaker) committed in a conversation with an agent $Ag_2$ (the addressee)
is a 7-tuple $C_{Ag_1,Ag_2} = \big<\mathcal{S},\mathcal{T},T,s,P_{Ag_1,Ag_2},CK\big>$ where:
\begin{itemize}
  \item $\mathcal{S}$ is the speaker's strategy;

  \item $\mathcal{T}$ is the current speaker's tactic, which we will define after introducing all the context elements;

  \item $T$ is a formula of $\pounds$ representing the negotiation topic that corresponds to the global goal;

  \item $s$ is a formula of $\pounds$ representing the argument on which the speaker should act;

  \item $P_{Ag_1,Ag_2}$ is the set of $Ag_{1}^{\prime}s$ beliefs about $Ag_{2}^{\prime}s$ beliefs $P^{bel}_{Ag_1,Ag_2}$ and
   about $Ag_{2}^{\prime}s$ preferences $P^{pref}_{Ag_1,Ag_2}$.   Thus $P_{Ag_1,Ag_2}$ = $P^{bel}_{Ag_1,Ag_2} \cup P^{pref}_{Ag_1,Ag_2}$;

  \item $CK$ is the common knowledge that the two agents share about the conversation.
\end{itemize}
In this definition we refine the context definition proposed in~\cite{Mbarki06} by adding the speaker strategy, the current
speaker tactic and the negotiation topic that corresponds to the global goal.
\end{definition}
During negotiation, $CK$ is constantly updated by adding all the information on which the two agents agree, including the accepted
arguments. We also assume that $CK \cap P_{Ag_1,Ag_2} = \emptyset$. In the context $C_{Ag_1,Ag_2}$, the influence of the
strategy on the tactic is reflected through the link between the topic $T$, the current argument $s$ and the strategy
$\mathcal{S}$. Let $\mathbb{T}$ be the set of topics ($T \in \mathbb{T}$), $g$ the current goal ($g \in \mathbb{B}$), $C$ the
sub-set of current constraints ($C \in \mathbb{C}$), $\gamma$ the sub-set of beliefs the agent is currently considering ($\gamma \in
\Gamma$), and $\mathbb{A}$ the set of arguments ($s \in \mathbb{A}$), we define the tactic as follows:

\begin{definition}{\emph{[Tactic]. }} \label{tactic}
A tactic is a function:
\begin{equation}\label{equation2}
\mathcal{T}: \mathbb{T} \times \mathcal{S}(g,C,\gamma) \times
\mathbb{A} \rightarrow \mathbb{B}
\end{equation}
\end{definition}

The key idea is that the current action (at the tactic level) is related to a sub-goal, which is determined by the strategy. From
the operational perspective, the current argument $s$ can attack or support the formula representing the sub-goal $T$. In order to
define the logical relation between $T$ and $s$, the authors in \cite{Mbarki06} introduced the notion of argumentation tree and
the notion of path that are defined as follow.

\begin{definition}{\emph{[Argumentation Tree].}} \label{tree}
Let $Ag$ be the set of participating agents and $Arg \subseteq \mathbb{A}$ be the set of arguments used by the agents in the
dialogue. An argumentation tree $Tr$ is a 2-tuple $Tr =\big<N,\rightarrow\big>$ where:

\begin{itemize}
  \item $N = \{(Ag_i, (H, h)) \mid Ag_i \in Ag, (H, h) \in Arg \}$ is the set of nodes. Each node is described as a pair $(Ag_i, (H, h))$,
  which indicates that the argument $(H, h)$ is used by the agent $Ag_i$.
  \item $ \rightarrow \subseteq  N \times N $ is a relation between nodes. We write $n_0 \rightarrow n_1$ instead of $(n_0, n_1) \in \rightarrow $
  where $\{n_0, n_1\} \subseteq N$. The relation $ \rightarrow $ is defined as follows:
       $(Ag_1, (H, h)) \rightarrow (Ag_2, (H^{\prime}, h^{\prime}))$ iff $Ag_1 \neq Ag_2$ and $(H^{\prime}, h^{\prime})$ attacks $(H, h)$.
\end{itemize}
\end{definition}

\begin{definition}{\emph{[Path]. }} \label{path}
Let $Tr = \big<N, \rightarrow \big>$ be an argumentation tree. A path in $Tr$ is a finite sequence of nodes $n_0, n_1,\cdots , n_m$
such that $\forall i$ ,$   0 \leq i < m $ : $n_i \rightarrow n_{i+1}$.
\end{definition}

In order to distinguish between relevant and irrelevant arguments in a given context  let us recall the definition of irrelevant
arguments:

\begin{definition}{\emph{[Irrelevant Argument]. }} \label{irrelevant}
Let $C_{Ag_1,Ag_2}$ = $\big<\mathcal{S}, \mathcal{T},T, s,P_{Ag_1,Ag_2},CK \big>$ be a conversation context, $Ag$ be the
set of participating agents, $Tr = \big<N, \rightarrow \big>$ be the argumentation tree associated to the conversation, and $(Ag_i,
(H, h))$ be a node in $Tr$ where $i \in \{1, 2\}$. $(H, h)$ is irrelevant in the context $C_{Ag_1,Ag_2}$ iff:

\begin{enumerate}
  \item  There is no path between the node $(Ag_i, (H, h))$ and the root of $Tr$ or;
  \item  $\exists x \in  CK : H \vdash \neg x$.

\end{enumerate}
\end{definition}

The distinction between relevant and irrelevant arguments allows agents to eliminate irrelevant arguments at each dialogue step before
ordering the relevant arguments in order to select the most relevant one.
Agents also have private preferences about different knowledge. Therefore, they may have private preferences about arguments.
This preference relation denoted by $(H, h)  \ll^{Ag_i}_{pref}  (H^{\prime}, h^{\prime})$ means that agent $Ag_i$ prefers the
argument $(H^{\prime}, h^{\prime})$ to the argument $(H, h)$.

\begin{definition}{\emph{[Preference]. }} \label{preference}
Let $(H, h)$ and $(H^{\prime}, h^{\prime})$ be two arguments. $(H, h)  \ll^{Ag_i}_{pref}  (H^{\prime}, h^{\prime})$ iff level $(H^{\prime}) \leq$ level $(H)$.
\end{definition}

Because $\leq$ is an ordering relation, the preference relation $\ll^{Ag_i}_{pref}$ is reflexive, antisymmetric, and transitive.
Agents may also have favorites among their arguments. How an agent favors an argument over others depends on the dialogue type. For
example, in a persuasive dialogue, an agent can favor arguments having more chances to be accepted by the addressee. In order to
characterize this notion, we introduce the notion of weight of an argument. The weight of an argument $(H,h)$ compared to another
argument $(H^{\prime},h^{\prime})$ in the context $C_{Ag_1,Ag_2}$ = $\big<\mathcal{S},\mathcal{T},T, s,P_{Ag_1,Ag_2},CK\big>$ is
denoted by $W_{(H,h)/(H^{\prime},h^{\prime})}^{P_{Ag_1,Ag_2}}$ and is evaluated according to the following algorithm:

\begin{algorithm}{\emph{\textbf{Algorithm} 1.}} \label{algorithm1}\\
\\ Step 1:  $ W_{(H,h)/(H^{\prime},h^{\prime})}^{P_{Ag_1,Ag_2}} = 0 . $ \\
\\Step 2:  $(\forall x \in H),(\forall x^{\prime}  \in H^{\prime} ):
 ( pref(x,x^{\prime}) \in P_{Ag_1,Ag_2}^{pref})\Rightarrow W_{(H,h)/(H^{\prime},h^{\prime})}^{P_{Ag_1,Ag_2}} = W_{(H,h)/(H^{\prime},h^{\prime})}^{P_{Ag_1,Ag_2}}  +  1 $
\end{algorithm}

$pref(x,x^{\prime}) \in P_{Ag_1,Ag_2}^{pref} $ means that $Ag_1$ believes that $Ag_2$ prefers $x$ to $x^{\prime}$. According to
this algorithm, the weight of an argument $(H, h)$ compared to another argument $(H^{\prime}, h^{\prime})$ is incremented by 1
each time $Ag_1$ believes that $Ag_2$ prefers a knowledge in $H$ to a knowledge in $H^{\prime}$. Indeed, each element of $H$ is
compered once to each element of $H^{\prime}$ according to the preference relation. Consequently, the computation algorithm of
the weight of an argument is finite because $H$ and $H^{\prime}$ are finite sets. The favorite relation is denoted by
$\preceq_{fav}^{P_{Ag_1,Ag_2}}$ and the strict favorite relation is denoted by $\prec_{fav}^{P_{Ag_1,Ag_2}}$.  $(H, h)
\preceq_{fav}^{P_{Ag_1,Ag_2}}  (H^{\prime},h^{\prime})$ means that agent $Ag_1$ favors the argument $(H^{\prime},h^{\prime})$ over
the argument $(H, h)$ according to $P_{Ag_1,Ag_2}$. This relation is defined as follows.

\begin{definition}{\emph{[Favorite Argument]. }} \label{favorite}
Let $C_{Ag_1,Ag_2}$ = $\big<\mathcal{S},\mathcal{T},T, s,P_{Ag_1,Ag_2},CK\big>$ be a conversation context and $(H, h)$
and $(H^{\prime},h^{\prime})$ be two arguments in the context $C_{Ag_1,Ag_2}$. We have:
\begin{itemize}
\item $(H, h)\preceq _{fav}^{P_{Ag_1,Ag_2}} (H^{\prime},h^{\prime})$ iff   $ W_{(H,h)/(H^{\prime},h^{\prime})}^{P_{Ag_1,Ag_2}} \leq W_{(H^{\prime},h^{\prime})/(H,h)}^{P_{Ag_1,Ag_2}}$ , \\
\item $(H, h)\preceq _{fav}^{P_{Ag_1,Ag_2}} (H^{\prime},h^{\prime})$ iff   $ W_{(H,h)/(H^{\prime},h^{\prime})}^{P_{Ag_1,Ag_2}} < W_{(H^{\prime},h^{\prime})/(H,h)}^{P_{Ag_1,Ag_2}}$.
\end{itemize}
\end{definition}

In order to allow agents to select the most relevant argument in a conversation context, an ordering relation between relevant
arguments is used. This ordering relation depends on the adopted strategy and is based on the notion of the risk of failure of an
argument. This notion of risk is subjective and there are several heuristics to evaluate it. In this paper, we use a heuristic based
on the fact that $CK$ contains certain knowledge and $P_{Ag_1,Ag_2}$ contains uncertain beliefs. To define this notion
formally, let $|H|_{\emptyset}$ be the number of formulas in $H$ that are not in $CK \cup P_{Ag_1,Ag_2}$, $|H|_{CK}$ be the number
of formulas in $H$ that are in $CK$, and $|H|_{P_{Ag_1,Ag_2}}$ be the number of formulas in $H$ that are in $P_{Ag_1,Ag_2}$.

\begin{definition}{\emph{[Risk of Failure of an Argument]. }} \label{risk}
Let $C_{Ag_1,Ag_2}$ = $\big<\mathcal{S},\mathcal{T},T, s,P_{Ag_1,Ag_2},CK\big>$ be a conversation context and $(H, h)$
and $(H', h')$ be two relevant arguments in the context $C_{Ag_1,Ag_2}$. The risk of failure $risk(H, h)$ of an argument
$(H, h)$ is a function mapping an argument to a natural number that satisfies the following property: $risk(H, h)\geq risk(H',
h')$ iff:
\begin{itemize}
\item   $|H|_{\emptyset} \geq |H'|_{\emptyset}$; or

\item   $|H|_{\emptyset} = |H'|_{\emptyset}$ and $|H|_{CK} \leq
|H'|_{CK}$; or

\item $|H|_{\emptyset} = |H'|_{\emptyset}$, $|H|_{CK} =
|H'|_{CK}$, and $|H|_{P_{Ag_1,Ag_2}} \leq |H'|_{P_{Ag_1,Ag_2}}$.
\end{itemize}
\end{definition}

Other approaches like those used in fuzzy systems to reason with uncertainty (using for example probabilities and possibilities)
can also be used to evaluate the risk of an argument. The advantages of our approach are its straightforward implementation
and additivity over uncertain hypotheses, which means adding uncertain hypotheses increases the risk of failure of an argument.
The relevance ordering relation denoted by $\preceq_r$ is defined as follows.

\begin{definition}{\emph{[Relevance Ordering Relation].}} \label{relevance}
Let $C_{Ag_1,Ag_2} = \big<\mathcal{S},\mathcal{T},T, s,P_{Ag_1,Ag_2},CK\big>$ be a conversation context and $(H, h)$
and $(H^{\prime},h^{\prime})$ be two relevant arguments in the context $C_{Ag_1,Ag_2}$. $(H^{\prime},h^{\prime})$ is more
relevant than $(H, h)$ denoted by $(H, h) \preceq_r (H^{\prime},h^{\prime})$ iff:
\begin{itemize}
\item $risk(H^{\prime},h^{\prime}) < risk(H, h)$; or

\item $risk(H^{\prime},h^{\prime}) = risk(H, h)$ and $(H, h) \prec_{fav}^{P_{Ag_1,Ag_2}} (H^{\prime},h^{\prime})$; or

\item $risk(H^{\prime},h^{\prime}) = risk(H, h)$ and $(H, h) \preceq_{fav}^{P_{Ag_1,Ag_2}} (H^{\prime},h^{\prime})$ and
$(H^{\prime},h^{\prime}) \preceq_{fav}^{P_{Ag_1,Ag_2}} (H,h)$ and $(H,h) \ll_{pref}^{Ag_1} (H^{\prime},h^{\prime}) $.
\end{itemize}
\end{definition}

According to this definition, $(H^{\prime},h^{\prime})$ is more relevant than $(H, h)$ if the risk of $(H, h)$ is greater than the
risk of $(H^{\prime},h^{\prime})$. If the two arguments have the same risk, the more relevant argument is the more favorable one
according to the favorite relation. If the two arguments have the same risk and they are equal according to the favorite relation,
the more relevant argument is the more preferable one according to the preference relation. The two arguments have the same relevance
if in addition they are equal according to the reference relation.

\subsubsection{Computational complexity of the Selection Mechanism}

Computationally speaking, the arguments selection mechanism is based on :
\begin{enumerate}
 \item The elimination of irrelevant arguments;
 \item The construction of new relevant arguments;
 \item The ordering of the relevant arguments using the relevance ordering relation; and
 \item The selection of one of the most relevant arguments.

\end{enumerate}

This process is executed by each participating agent at each dialogue step at the tactical level. The relevant arguments that
are not selected at a step $t_i$, are recorded and added to the set of potential arguments $PA$ because they can be used at a
subsequent step. The set of potential arguments can be viewed as a stack in which the higher level argument is the most relevant one.
A relevant argument constructed at a step $t_i$ and used latter at a step $t_j$ simulates the backtracking towards a previous node in
the argumentation tree and the construction of a new path.

Here we prove that our selection mechanism is tractable if arguments are represented using propositional Horn clauses.
propositional Horn clauses is a restricted language that has been proved to be sufficient to represent and reason about
knowledge in many concrete applications \cite{BentaharIEEEIS2007}. A propositional Horn clause is a disjunction of literals, which are
atomic propositions (called positive literals) or their negations (called negative literals), with at most one positive literal.
Formally, a propositional Horn clause has the form $\neg p_1 \vee \neg p_2 \vee \dots \vee \neg p_n \vee c$ or also $p_1 \wedge p_2
\wedge \dots \wedge p_n \rightarrow c$, which is simply an implication. A propositional Horn formula is a conjunction of
propositional Horn clauses. We focus on a further restriction called propositional definite Horn clauses, where each clause has
exactly one positive literal. A propositional definite Horn formula is a conjunction of propositional definite Horn clauses.
This restriction is of particular interest in modeling argumentation reasoning for negotiation, because formulas of the
type $p_1 \wedge p_2 \wedge \dots p_n \rightarrow c$ are adequate to describe interrelationships between premises (i.e., reasons or
justifications) and conclusions (i.e., offers). Thus, agents could support their offers (the part $c$) as positive literals using the
support $p_1 \wedge p_2 \wedge \dots p_n$.

\begin{theorem}\label{Complexity1}
If arguments are represented in proportional definite Horn clauses, the arguments selection mechanism runs in polynomial time.
\end{theorem}

\noindent\emph{\textbf{Proof.}} \emph{It is known from Bentahar et al. \cite{BentaharIEEEIS2007}, that given a Horn knowledge base
$\Gamma$, a subset $H \subseteq \Gamma$, and a formula $h$; checking whether $(H, h)$ is an argument is polynomial. To decide
if an argument is irrelevant, we have to check if 1) $H \vdash \neg x$ for an $x \in CK$, which can be done in polynomial time
since $H$ is a definite Horn formula; or 2) there is a path from the root to $(H, h)$, which is a graph reachability problem, and
it is known by Jones \cite{Jones} that the problem is in NLOGSPACE. Since NLOGSPACE $\subseteq$ P, the problem can be
solved in polynomial time. To decide about the preference, we only need to compute the level of an argument from the level of a
subset of $\Gamma$, which is  a simple procedure that is obviously polynomial. Computing the favorite argument given two arguments
needs the computation of the arguments' weight, which is again a polynomial procedure as shown by Algorithm 1. To compere two given
arguments using the risk, we only need to compute the number of formulas in $H$ and check if they are part of different sets,
which is a polynomial procedure. Finally, the relevance ordering relation is simply based on comparing risks and favorites, which
are both polynomial, so we are done.}~~$\blacksquare$
 \section{Agent's Uncertainty}\label{sec:uncertainty}

\subsection{Generalities and Overview}

The process of selecting arguments is always associated with a degree of uncertainty. Despite the agent will select the most relevant argument
at each dialogue step using his theory and tactic reasoning, however, there still be a doubt that the selected argument is the right one
and it will be accepted by the addressee.  To measure this uncertainty, we use Shannon entropy as we did in our pervious work ~\cite{MareyBE09}, a
well-known technique that defines and quantifies the information. The idea behind Shannon entropy in information theory is based on the amount of
randomness that exists in a random event. In dialogue games, we assume that at each dialogue step, the agent has different choices of arguments
(i.e., the set of potential arguments $PA$ at that step) and he will select one of them. The selection is based on the speaker's knowledge base and
characterized by an amount of uncertainty over this base and an amount of randomness over the addressee's knowledge, beliefs, and preferences. Indeed,
Shannon entropy is a measure of the uncertainty associated with a random variable (i.e., in our case, the selected argument). The more uncertain we are
about the content of the message, the more informative it is~\cite{CoverT91}.

\begin{definition}{\emph{[Shannon entropy]. }} \label{entropy}
Shannon entropy for a discrete random variable \emph{$X$} taking its values from a set of values \emph{$S$} (sample space),
with probability mass function \emph{$P(x)$} is given by Equation~\ref{equation1}:

\begin{equation}\label{equation1}
H(X) = - \sum_{x\in S} P(x)Log P(x)
\end{equation}

\end{definition}

Shannon entropy $H(X)$ depends on the probability distribution of $X$ rather than the actual values of $X$. The logarithm in
Equation~\ref{equation1} is considered to be of base 2 in the computations. The value of $H(X)$ varies from zero to $Log(|S|)$,
where zero means that there is no uncertainty, while $Log(|S|)$ is the maximum value of uncertainty. In this paper, we aim to
investigate to what extent we can use Shannon entropy in the evaluation of the agent's uncertainty in the dialogue. We place
ourselves in the role of an external observer trying to evaluate the uncertainty of the participants to the dialogue i.e., how
certain/uncertain each agent is about the selected move, how certain/uncertain he is that the selected move will be accepted by
the addressee and the distinction between these tow types of uncertainty. The basic idea is to measure how much the agent is
uncertain about selecting the right move $M_i$ at each step $t_i$ in dialogue $D$ by assuming that there is a set $S_i$ of choices
facing the agent at each dialogue step. Throughout the paper, we will measure the agent's uncertainty using Shannon entropy, then
we normalize it to have a value between zero and one and subtracting it from one to get the agent's certainty for selecting the right
move at that step. We call this measure the \emph{certainty index} ``$CI(M_i)$" of that move, and then we calculate how much the agents
are certain about the whole dialogue using what we call the \emph{certainty index of the dialogue} ``$CI(D)$" by taking the average of
the certainty index of all moves in this dialogue (taking the minimum is another choice that we also discuss).
Furthermore, we measure the certainty index of the whole dialogue by computing all possible dialogues, using the Cartesian product
of all possible moves, and determining the probability of each dialogue, and then applying the general formula of Shannon entropy
for the whole dialogue (exactly as what we do in the case of calculating the certainty of the moves). Moreover, we differentiate
between to types of uncertainty; namely Type 1, and Type 2, and for the first time we classify the arguments into three classes based on
three criteria that we will introduce later on.

To allow agents to refer to their dialogue history, a data structure called commitment store $``CS"$ is used to restore
utterances that agents utter during the dialogue~\cite{Hamblin70}.
Let $Ag_1$ and $Ag_2$ be two agents $Ag_1 \neq Ag_2$. Also, let $\Gamma_{Ag_x}$ be $Ag_x's$ knowledge base ($x \in \{1,2\}$).
$CS_{Ag_x}^{t_i}$ is the commitment store of agent $Ag_x$ at step $t_i$ of the dialogue. Suppose that at step $t_{i-1}$, agent
$Ag_2$ uttered a move. To utter a move at the next step $t_i$, agent $Ag_1$ should consider his knowledge base and the content of
$Ag_2's$ commitment store. Let $m_{i}^j$ be the $j^{th}$ move among the possible moves an agent has at the step $t_i$ and
$P(m_{i}^j)$ the associated probability such that the relationship between the move $M_i$ and $m_{i}^j$ is as follows:

\begin{equation}\label{equation3}
\forall i ~ \exists j : M_{i} = m_{i}^j
\end{equation}

where $M_{i}$ is the selected move the agent utters at the step $t_{i}$, and the production of moves for $Ag_1$ along with their
probabilities is a function of $\Gamma_{Ag_1}$ and $CS_{Ag_2}^{t_i}$:

\begin{equation}\label{equation4}
\zeta(\Gamma_ {Ag_1} \cup  CS_{Ag_2}^{t_i}) = \{(m_{i}^j , P(m_{i}^j)) | m_{i}^j\in S_{i}\}
\end{equation}
where $S_i$ is the set of choices facing the agent at the dialogue step $t_i$.

We belief that such measures can help the participants to the dialogue make a better decisions for playing the most appropriate argument at
each moment (i.e., at the tactical level) to achieve their agreements based on the adapted strategy, and it will help evaluate dialogues and strategies
of the participants to these dialogues. In what follows we introduce the two types of agent's uncertainty.

%In~\cite{MareyBE09}, we proposed a set of metrics using Shannon entropy. We placed ourselves in the role of an external observer
%to evaluate the dialogue. The basic idea is to measure how much the agents are uncertain/certian about selecting the right move
%$M_i$ at each step $t_i$ in dialogue $D$ by assuming that there is a set of potential moves (i.e., arguments) facing the agent at
%each dialogue step. Based on that, we can calculate the certainty index of the dialogue $CI(D)$, which reflects the certainty degree
%of the agents about the whole dialogue. Such measures will help evaluate dialogues and strategies of the
%participants to these dialogues.
%
%In this paper, we extend our previous work to evaluate the uncertainty/certainty index that the selected move will be
%accepted by the addressee and analyze the relationship between these two types of uncertainty (i.e., the agent's uncertainty about
%selecting the right move and the agent's uncertainty that the selected move will be accepted by the addressee). Moreover, we
%analyze special cases in which we cannot consider the certainty index as an indicator of the agent's certainty that the selected
%move will be accepted by the addressee, but as a representation of the agent's certainty about selecting the right move at that step.
%For instance, when there is only one argument, or all the available arguments have the same probability. We belief that such
%measures can help the participants to the dialogue make decisions for playing the most appropriate argument at each moment (at the
%tactical level) to achieve their goals based on the adapted strategy.

%For simplification reasons, we will consider only two agents involved in the dialogue $Ag_1$ and $Ag_2$. Multi-party dialogue
%is an important issue that we plan to investigate in future work. During a dialogue, several moves may be uttered. Those moves
%constitute a dialogue $D$, which is a sequence of moves denoted by $[M_0,M_1,\cdots ,M_n]$, where $M_0$ is the initial move, $M_n$ is
%the final one, and $|D|$ is the length of the dialogue (i.e., $|D|
%= n$).

\subsection{Agent's Uncertainty: Type I}\label{sec:type1}

In this section, we will discuss to what extent we can use Shannon entropy in dialogue games to measure the uncertainty/certainty
index of the agents about their dialogue. To do so, we measure the uncertainty index of the agents about their moves and then we
compute the uncertainty index about the whole dialogue. The following two subsections illustrate these measures.

\subsubsection{Measuring How Certain the Agent is about his Move}\label{moveuncertainty}
To measure the uncertainty index of the agent about his move, we calculate Shannon entropy of that move, which means the agent's
uncertainty about selecting the right move at a given dialogue step. Here we consider only possible moves, which are moves whose
the associated probability is in $[0,1]$. Then we normalize this value by dividing it by the logarithm of the number of all
possible moves at that step. By subtracting the normalized value from one; we get the agent's certainty index about selecting the
right move at this step.

\begin{definition}{\emph{[Move's entropy]. }} \label{Move'sentropy}
Let $D = [M_0, M_1, \ldots, M_n]$ be a negotiation dialogue, and suppose that at each dialogue step $t_i$ a set $S_i$ of moves
$\{m_i^1,m_i^2,\ldots,m_i^k\}$ are possible, and each one of them is associated with a given probability $P(m_i^j) = 1-risk(m_i^j)$,
such that $\sum_{m_i^j\in S_i} P(m_i^j)=1$. Shannon entropy for a random move $M_i$ taking its values from the set of moves $S_i$ is
defined by:
\begin{equation}\label{equation5}
H(M_i) = - \sum_{m_i^j\in S_i}P(m_i^j)Log P(m_i^j)
\end{equation}
\end{definition}

The value of $H(M_i)$ varies from zero to $Log(|S_i|)$, where zero means that there is no uncertainty (i.e., there is only one
choice), while $Log(|S_i|)$ means that the uncertainty is at its maximum value (i.e., all moves have the same probability). We
further normalize $H(M_i)$ to have a metric that ranges from 0 to 1. This can be achieved by dividing $H(M_i)$ by $Log(|S_i|)$. So,
the uncertainty about selecting the right move is given by:

\begin{equation}\label{equation6}
\mu(M_i)=
\begin{cases} 0 & \text{\emph{iff}}~~~Log(|S_i|)=0
\\
H(M_i)/Log(|S_i|) &otherwise
\end{cases}
\end{equation}

\begin{proposition} \label{proposition1}
The uncertainty of a move $M_i$ at a certain step $t_i$ during the dialogue is equal to zero (i.e., $\mu(M_i)=0$) iff at that step the
agent has only one choice.
\end{proposition}

\noindent\emph{\textbf{Proof.}}\\
\emph{$\mu(M_i)= 0 \Leftrightarrow Log(|S_i|)=0$\\
\\ $\Leftrightarrow  |S_i|=1$.~~$\blacksquare$}

>>>>

\setcounter{chapter}{2}
\chapter{Contribution and Research Activities}

%******************** START Contribution and Research Activities. ***************
In this chapter we will introduce the important notions in argumentation-based negotiation,
\section{Argumentation System and Agent Theory}\label{sec:Argumentation}
